Modern BTree techniques

### 2022-09-22 strangeloop/presentation.tex
Adjustments

    \begin{itemize}[label={\MVRightarrow}]
        \item RUM conjecture
        \item B-Tree internals
        \begin{itemize}[label={}]
            \item Key normalization
            \item Prefix/Suffix truncation
            \item SB-Tree
        \end{itemize}
        \item B-Tree as a component
        \begin{itemize}[label={}]
            \item Hybrid index
            \item Bw-Tree
            \item Partitioned B-Tree
        \end{itemize}
        \item Learned Indexes
        \item Everything else
    \end{itemize}
\end{frame}

@@

    \begin{center}
    \textbf{SB-tree}
    \vspace{1cm}

    \begin{overprint}[15cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \btreenode{1}{}{btree-hide}{btree-branch}{}
            \btreenode{2}{right=2cm of node1.east}{btree-hide}{btree-branch}{}
            \node[btree-hide, right=2cm of node2.east, minimum width=0.5cm] (node2-extra) {};

            \btreenode{3}{below=1cm of node1.south west, xshift=-0.5cm}{btree-hide}{btree-leaf}{}
            \btreenode{4}{below=1cm of node1.south east, xshift=0.5cm}{btree-hide}{btree-leaf}{}
            \btreenode{5}{right=2cm of node4.east}{btree-hide}{btree-leaf}{}
            \node[btree-hide, right=2cm of node5.east, minimum width=0.5cm] (node5-extra) {};

            \draw[->, btree-line] (pointer21.south)
                .. controls ([yshift=-1cm] pointer21) and ([yshift=1cm] pointer14) ..
                (pointer14.north);
            \draw[->, btree-line] (pointer31.south)
                .. controls ([yshift=-1cm] pointer31) and ([yshift=1cm] pointer13) ..
                (pointer13.north);

            \node[box, inner sep=0.2cm,
                fit=(node1)(node2)(node2-extra)
            ] (extent-1) {};

            \node[box, inner sep=0.2cm,
                fit=(node3)(node4)(node5)(node5-extra)
            ] (extent-2) {};
        \end{tikzpicture}

        \onslide<2>
        \begin{tikzpicture}[]
            \btreenode{1}{}{btree-hide}{btree-branch}{}
            \btreenode{2}{right=2cm of node1.east}{btree-hide}{btree-branch}{}
            \node[btree-hide, right=2cm of node2.east, minimum width=0.5cm] (node2-extra) {};

            \btreenode{3}{below=1cm of node1.south west, xshift=-0.5cm}{btree-hide}{btree-leaf}{}
            \btreenode{4}{below=1cm of node1.south east, xshift=0.5cm}{btree-hide}{btree-leaf}{}
            \btreenode{5}{right=2cm of node4.east}{btree-hide}{btree-leaf}{}
            \node[btree-hide, right=2cm of node5.east, minimum width=0.5cm] (node5-extra) {};

            \draw[->, btree-line] (pointer21.south)
                .. controls ([yshift=-1cm] pointer21) and ([yshift=1cm] pointer14) ..
                (pointer14.north);
            \draw[->, btree-line] (pointer31.south)
                .. controls ([yshift=-1cm] pointer31) and ([yshift=1cm] pointer13) ..
                (pointer13.north);

            \node[box, inner sep=0.2cm,
                fit=(node1)(node2)(node2-extra)
            ] (extent-1) {};

            \node[box, inner sep=0.2cm,
                fit=(node3)(node4)(node5)(node5-extra)
            ] (extent-2) {};

            \draw[->, btree-line] (node5.center)
                .. controls ([yshift=-1cm] node5) and ([yshift=-1cm,xshift=-1cm] node5-extra) ..
                ([xshift=-1cm]node5-extra.center);
        \end{tikzpicture}
    \end{overprint}

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        O’Neil P. E. (1992). The SB-tree: An index-sequential structure for high performance sequential access. \\
        Acta Informatica.
    \linespread{1.5}

    \end{center}
\end{frame}
\note{
    OrientDB uses SB-Tree, but are they the same?
}

@@

    \begin{center}
        \begin{itemize}[label={\MVRightarrow}]
            \item <+-> Buffered (Lazy) B-Tree
            \item <+-> Cache conscious B-Tree
            \item <+-> Distributed B-Tree
            \item <+-> UB-Tree
            \item <+-> MDAM
            \item <+-> LSM Tree
            \item <+-> Trie
            \item <+-> ...
        \end{itemize}
    \end{center}
\end{frame}
\note{
    Buffered B-Tree are used in WiredTiger.
    Skip list are used also in WiredTiger as in memory structure, in Apache
    Cassandra again as in memory, and in Redis.
}

>>>>

    \begin{itemize}[label={\MVRightarrow}]
        \item RUM conjecture
        \item B-Tree internals
        \begin{itemize}[label={}]
            \item \footnotesize{Key normalization}
            \item \footnotesize{Prefix/Suffix truncation}
            %\item SB-Tree
        \end{itemize}
        \item B-Tree as a component
        \begin{itemize}[label={}]
            \item \footnotesize{Hybrid index}
            \item \footnotesize{Bw-Tree}
            \item \footnotesize{Partitioned B-Tree}
        \end{itemize}
        \item Learned Indexes
        \item Everything else
    \end{itemize}
\end{frame}
\note{
    small sub points
}

@@

    %\begin{center}
    %\textbf{SB-tree}
    %\vspace{1cm}

    %\begin{overprint}[15cm]
        %\onslide<1>
        %\begin{tikzpicture}[]
            %\btreenode{1}{}{btree-hide}{btree-branch}{}
            %\btreenode{2}{right=2cm of node1.east}{btree-hide}{btree-branch}{}
            %\node[btree-hide, right=2cm of node2.east, minimum width=0.5cm] (node2-extra) {};

            %\btreenode{3}{below=1cm of node1.south west, xshift=-0.5cm}{btree-hide}{btree-leaf}{}
            %\btreenode{4}{below=1cm of node1.south east, xshift=0.5cm}{btree-hide}{btree-leaf}{}
            %\btreenode{5}{right=2cm of node4.east}{btree-hide}{btree-leaf}{}
            %\node[btree-hide, right=2cm of node5.east, minimum width=0.5cm] (node5-extra) {};

            %\draw[->, btree-line] (pointer21.south)
                %.. controls ([yshift=-1cm] pointer21) and ([yshift=1cm] pointer14) ..
                %(pointer14.north);
            %\draw[->, btree-line] (pointer31.south)
                %.. controls ([yshift=-1cm] pointer31) and ([yshift=1cm] pointer13) ..
                %(pointer13.north);

            %\node[box, inner sep=0.2cm,
                %fit=(node1)(node2)(node2-extra)
            %] (extent-1) {};

            %\node[box, inner sep=0.2cm,
                %fit=(node3)(node4)(node5)(node5-extra)
            %] (extent-2) {};
        %\end{tikzpicture}

        %\onslide<2>
        %\begin{tikzpicture}[]
            %\btreenode{1}{}{btree-hide}{btree-branch}{}
            %\btreenode{2}{right=2cm of node1.east}{btree-hide}{btree-branch}{}
            %\node[btree-hide, right=2cm of node2.east, minimum width=0.5cm] (node2-extra) {};

            %\btreenode{3}{below=1cm of node1.south west, xshift=-0.5cm}{btree-hide}{btree-leaf}{}
            %\btreenode{4}{below=1cm of node1.south east, xshift=0.5cm}{btree-hide}{btree-leaf}{}
            %\btreenode{5}{right=2cm of node4.east}{btree-hide}{btree-leaf}{}
            %\node[btree-hide, right=2cm of node5.east, minimum width=0.5cm] (node5-extra) {};

            %\draw[->, btree-line] (pointer21.south)
                %.. controls ([yshift=-1cm] pointer21) and ([yshift=1cm] pointer14) ..
                %(pointer14.north);
            %\draw[->, btree-line] (pointer31.south)
                %.. controls ([yshift=-1cm] pointer31) and ([yshift=1cm] pointer13) ..
                %(pointer13.north);

            %\node[box, inner sep=0.2cm,
                %fit=(node1)(node2)(node2-extra)
            %] (extent-1) {};

            %\node[box, inner sep=0.2cm,
                %fit=(node3)(node4)(node5)(node5-extra)
            %] (extent-2) {};

            %\draw[->, btree-line] (node5.center)
                %.. controls ([yshift=-1cm] node5) and ([yshift=-1cm,xshift=-1cm] node5-extra) ..
                %([xshift=-1cm]node5-extra.center);
        %\end{tikzpicture}
    %\end{overprint}

    %\linespread{0.5}
    %\vspace{0.5cm}
    %\color{black}\fontsize{6pt}{0}\selectfont
        %O’Neil P. E. (1992). The SB-tree: An index-sequential structure for high performance sequential access. \\
        %Acta Informatica.
    %\linespread{1.5}

    %\end{center}
%\end{frame}
%\note{
    %OrientDB uses SB-Tree, but are they the same?
%}

@@

    \begin{center}
        \begin{itemize}[label={\MVRightarrow}]
            \item Buffered (Lazy) B-Tree
            \item Cache conscious B-Tree
            \item Distributed B-Tree
            \item SB-Tree
            \item MDAM
            \item LSM Tree
            \item Trie
            \item ...
        \end{itemize}
    \end{center}
\end{frame}
\note{
    Buffered B-Tree are used in WiredTiger.
    Skip list are used also in WiredTiger as in memory structure, in Apache
    Cassandra again as in memory, and in Redis.
}
